---
title: "IDS investigation worksheet"
author: "by Team-Name: User1, User2, User3, User4 & User5"
date: "`r Sys.Date()`"
output: html_document
---

**Note:** You can use this file as you 'working document' where you can try out various investigation ideas and keep notes about your findings. How you use and structure this file is up to you. It is recommended that you keep notes about what you are investigating and what you find as this will make the process of creating your presentation and report easier. Please note that you _do not_ need to submit this file as part of your group project.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-lib, message = FALSE}
library(tidyverse)
library(Metrics)
library(ggplot2)
```


```{r load-data}
# load your data 
laptop_data <- read.csv("~/final project/data/laptop_data_cleaned.csv")

# Preview the data
head(laptop_data)

# Check for missing values
colSums(is.na(laptop_data))

# Remove missing values
laptop_data <- laptop_data %>% drop_na()

# Ensure numeric columns are correctly typed
laptop_data$Price <- as.numeric(laptop_data$Price)
laptop_data$Ram <- as.numeric(laptop_data$Ram)
laptop_data$Weight <- as.numeric(laptop_data$Weight)
laptop_data$Ppi <- as.numeric(laptop_data$Ppi)

# Convert categorical variables to factors
laptop_data$TypeName <- as.factor(laptop_data$TypeName)
laptop_data$Cpu_brand <- as.factor(laptop_data$Cpu_brand)

# Descriptive statistics
summary(laptop_data)

# Correlation matrix
cor_matrix <- laptop_data %>%
  select(Price, Ram, Weight, Ppi) %>%
  cor()
print(cor_matrix)

# Price distribution plot
ggplot(laptop_data, aes(x = Price)) +
  geom_histogram(fill = "blue", bins = 30, color = "black") +
  labs(title = "Price Distribution", x = "Price", y = "Frequency")

# RAM vs Price
ggplot(laptop_data, aes(x = Ram, y = Price, color = TypeName)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "RAM vs Price", x = "RAM (GB)", y = "Price")

# Price by CPU Brand
ggplot(laptop_data, aes(x = Cpu_brand, y = Price, fill = Cpu_brand)) +
  geom_boxplot() +
  labs(title = "Price by CPU Brand", x = "CPU Brand", y = "Price")

# One-hot encoding
laptop_data <- laptop_data %>%
  mutate(across(where(is.factor), ~ as.numeric(as.factor(.)) - 1))

# Split into training and testing datasets
set.seed(123)
train_index <- sample(1:nrow(laptop_data), 0.8 * nrow(laptop_data))
train_data <- laptop_data[train_index, ]
test_data <- laptop_data[-train_index, ]

# Extract features and target
X_train <- train_data %>% select(-Price)
y_train <- train_data$Price
X_test <- test_data %>% select(-Price)
y_test <- test_data$Price

# Train linear regression model
linear_model <- lm(Price ~ ., data = train_data)

# Model evaluation
summary(linear_model)

# Predict on test data
y_pred_linear <- predict(linear_model, newdata = test_data)

# Calculate RMSE
rmse_linear <- rmse(y_test, y_pred_linear)
cat("Linear Regression RMSE:", rmse_linear, "\n")

# Actual vs Predicted (Linear Regression)
ggplot(data.frame(Actual = y_test, Predicted = y_pred_linear), aes(x = Actual, y = Predicted)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(title = "Actual vs Predicted (Linear Regression)", x = "Actual Price", y = "Predicted Price")

# Comment on model
plot(linear_model, which = 1)
plot(linear_model, which = 2)
plot(linear_model, which = 3)
plot(linear_model, which = 4)
plot(linear_model, which = 5)
```


