---
title: "IDS investigation worksheet"
author: "by Team-Name: User1, User2, User3, User4 & User5"
date: "`r Sys.Date()`"
output: html_document
---

**Note:** You can use this file as you 'working document' where you can try out various investigation ideas and keep notes about your findings. How you use and structure this file is up to you. It is recommended that you keep notes about what you are investigating and what you find as this will make the process of creating your presentation and report easier. Please note that you _do not_ need to submit this file as part of your group project.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-lib, message = FALSE}
library(tidyverse)
library(Metrics)
library(ggplot2)
```


```{r load-data}
# Load your data
laptop_data <- read.csv("~/final project/data/laptop_data_cleaned.csv")

# Preview the data
head(laptop_data)

# Check for missing values
colSums(is.na(laptop_data))

# Remove missing values
laptop_data <- laptop_data %>% drop_na()

# Descriptive statistics
summary(laptop_data)

# Correlation matrix for numeric variables
cor_matrix <- laptop_data %>%
  select(Price, SSD, Weight, Ppi, Ram) %>%
  cor()
print(cor_matrix)

# Price distribution plot
ggplot(laptop_data, aes(x = Price)) +
  geom_histogram(fill = "blue", bins = 30, color = "black") +
  labs(title = "Price Distribution", x = "Price", y = "Frequency")

# RAM vs Price by TypeName
ggplot(laptop_data, aes(x = Ram, y = Price, color = TypeName)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "RAM vs Price", x = "RAM (GB)", y = "Price")

# Price by CPU Brand
ggplot(laptop_data, aes(x = Cpu_brand, y = Price, fill = Cpu_brand)) +
  geom_boxplot() +
  labs(title = "Price by CPU Brand", x = "CPU Brand", y = "Price")

# Split into training and testing datasets
set.seed(123)
train_index <- sample(1:nrow(laptop_data), 0.8 * nrow(laptop_data))
train_data <- laptop_data[train_index, ]
test_data <- laptop_data[-train_index, ]

# Train improved linear regression model
linear_model <- lm(
  Price ~ Weight + Ppi + SSD + Cpu_brand + TypeName + Ram,
  data = train_data
)

# Model evaluation
summary(linear_model)

# Predict on test data
y_pred <- predict(linear_model, newdata = test_data)

# Calculate RMSE
rmse_value <- rmse(test_data$Price, y_pred)
cat("Final Model RMSE:", rmse_value, "\n")

# Actual vs Predicted Plot
ggplot(data.frame(Actual = test_data$Price, Predicted = y_pred), aes(x = Actual, y = Predicted)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(title = "Actual vs Predicted (Final Model)", x = "Actual Price", y = "Predicted Price")

# Residual diagnostics
par(mfrow = c(2, 2))  # Arrange plots in 2x2 grid
plot(linear_model)

```


