---
title: "Predictive Analytics for Laptop Prices"
author: "by Team-Name: Zhenye Cao, Siran Luo, Yunxi Guan,Zongyang Jin, Tianmu Xu , Qinhao Feng(Shawn Feng)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-lib, include = FALSE}
library(tidyverse)
library(Metrics)
library(ggplot2)



```


```{r load-data, include=FALSE}
# Load your data
laptop_data <- read.csv("~/final-project/data/laptop_data_cleaned.csv")

# Preview the data
head(laptop_data)

# Check for missing values
colSums(is.na(laptop_data))

# Remove missing values
laptop_data <- laptop_data %>% drop_na()

# Descriptive statistics
summary(laptop_data)

# Correlation matrix for numeric variables
cor_matrix <- laptop_data %>%
  select(Price, SSD, Weight, Ppi, Ram) %>%
  cor()
print(cor_matrix)

# Split into training and testing datasets
set.seed(123)
train_index <- sample(1:nrow(laptop_data), 0.8 * nrow(laptop_data))
train_data <- laptop_data[train_index, ]
test_data <- laptop_data[-train_index, ]

# Train improved linear regression model
linear_model <- lm(
  Price ~ Weight + Ppi + SSD + Cpu_brand + TypeName + Ram,
  data = train_data
)
linear_model_2 <- lm(Price ~ ., data = train_data)

# Model evaluation
summary(linear_model)
summary(linear_model_2)

# Predict on test data
y_pred <- predict(linear_model, newdata = test_data)

# Calculate RMSE
rmse_value <- rmse(test_data$Price, y_pred)
cat("Linear Model RMSE:", rmse_value, "\n")

# Actual vs Predicted Plot
ggplot(data.frame(Actual = test_data$Price, Predicted = y_pred), aes(x = Actual, y = Predicted)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(title = "Actual vs Predicted (Linear Model)", x = "Actual Price", y = "Predicted Price")

# Residual diagnostics
par(mfrow = c(2, 2))  
plot(linear_model)



```


## Research Question

The objective of our project is to investigate how various factors influence laptop prices. We cleaned the data and utilized methods such as descriptive analysis, data visualization, and modeling to identify the variables affecting the price. 

The analysis employs tools like correlation matrices, linear regression, and visualization techniques to gain insights and build a predictive model. The outcomes include pricing trends, identification of significant predictors of price, and an evaluation of the predictive model. These results can inform decisions in the laptop market and guide both consumers effectively.

## Data
Data Description

First, Company, which includes popular brands like Apple and Dell. The TypeName indicates the different type of the laptop, such as gaming laptops or ultrabooks.RAM represents the temporary storage used for active data, which impacts performance and multitasking capabilities.Weight reflects the portability of the laptop, and Price is.

Additionally, we have Touchscreen (1 for yes, 0 for no) and IPS display (1 for yes, 0 for no).PPI (Pixels Per Inch) measures screen sharpness—the higher the PPI, the clearer the image. The CPU brand, like Intel Core i7, while HDD and SSD refer to storage capacity and speed.

Finally, the GPU brand impacts graphics performance, and the OS mostly indicates Windows as the operating system.

Model Building

In this model-building phase, our objective was to predict the price of laptops (Price) based on various factors. Initially, we explored the dataset by checking for missing values, applying transformations, and then used a linear regression model to assess the impact of each predictor on the target variable (Price). The model was subsequently refined to ensure its validity and relevance.

We first build a model with all variables, to analyze the influence of each variable. After constructing the initial model, we examined the significance (p-values) of each predictor using T-tests. T-tests helped identify variables that were individually significant in predicting the target variable (Price). For example, Ram, Ppi, and SSD showed low p-values, indicating their strong individual effects on Price. Conversely, some variables had high p-values, suggesting they were not statistically significant.

Using p-values from T-tests, we systematically removed insignificant variables from the model. For example, variables such as TouchScreen were excluded due to their high p-values (greater than 0.05), indicating their minimal contribution to predicting the target variable. This stepwise process ensured that only variables with significant individual contributions remained in the model. Additionally, the large number of companies made the model more complex, potentially leading to biases in data interpretation. To focus on the relationship between price and hardware characteristics, we decided to remove the Company variable in the new model, thereby disregarding the impact of brand.

The final model includes the following predictors: Ram, Weight, Ppi, SSD, Cpu_brand, TypeName, and their interaction terms. We assessed the model's fit using the adjusted R-squared (Adjusted R²). The final model achieved an R² of 0.8024, indicating that it explains 80.24% of the variability in laptop prices. Additionally, residuals were checked to ensure that the basic assumptions of linear regression were met: linearity, homoscedasticity, and normality.

In terms of prediction, the model showed minimal differences between predicted and actual prices, demonstrating good predictive performance. The Root Mean Squared Error (RMSE) was 0.2995415, suggesting that the model performs well on the test dataset. We also build the Actual and Predicted plot to check the model, it compares the true laptop prices with the prices predicted by the linear regression model, which helps us to assess the model’s predictive performance. The majority of points close to the diagonal line, indicating that the model performs well for most predictions

Model Checking

This section aims to check the assumptions for our final model. 

Residuals vs Fitted:
The residuals are mostly scattered around zero, suggesting that the model captures much of the variation. And there are no apparent systematic patterns. Therefore, the linearity assumption is not violated. However, the red smoothing line shows a slight curvature at the end of the plot, so there might have some evidence against the linearity assumption. The residuals have a roughly constant spread across all fitted values, which indicates the homoscedasticity. There seems to be outliers far from the majority of residuals, which may influence the model, so use Cook’s Distance to assess their impact later.

Q-Q Plot of Residuals:
The residuals align well with the diagonal line in the middle, confirming that the majority follow a normal distribution. Although a slight light-tailed distribution may be detected, points fit the line adequately. While this may not significantly impact the model's predictive capability, transforming the response variable could improve normality if required for inference.

Scale-Location Plot:
The plot shows a random scatter of points around the horizontal line. This suggests that the residuals have constant variance across the range of fitted values. By contrast, the spread of residuals increases slightly with fitted values, as evidenced by the upward trend in the red line. This indicates mild heteroscedasticity. While minor, applying weighted least squares regression or response transformations may improve the model's performance.

Residuals vs Leverage:
Points 981, 982, and 210 exhibit high leverage and fall near the Cook's Distance threshold, making them influential observations. However, because of the uniqueness of each observation (i.e. different properties of computers might have different influence on price), dropping an observation will mean that we don’t have specific data representing a particular property, causing our model unable to produce valid estimation of the impact of properties on computer price. Therefore, we decided not to drop any observation.

## Findings
The regression analysis reveals key factors influencing laptop prices. Among continuous variables, RAM has the strongest positive impact on price, with a coefficient of 0.0251, indicating that increased memory significantly raises laptop prices. PPI, which measures display quality, also has a positive influence (0.00269), with higher resolution displays leading to higher prices. Weight, however, has a smaller effect (0.0514), suggesting that lighter laptops are not necessarily more expensive.

For categorical variables, CPU type significantly affects pricing. Laptops with Intel Core i7 processors (0.5182) are the most expensive, followed by those with Intel Core i5 (0.4608) and Intel Core i3 (0.1492), while "other Intel" processors (-0.2132) reduce prices. Laptop type also plays a role, with workstations (0.5627) and gaming laptops (0.1297) commanding higher prices, whereas netbooks (-0.0796) and notebooks (-0.1515) are cheaper. Storage type contributes as well, with SSDs (0.0006718) adding a premium due to their faster performance.

Brand is another critical factor. High-end brands like Google (0.739), LG (0.679), and Apple (0.555) significantly increase prices due to their premium positioning. Mid-range brands, such as Dell (0.353), HP (0.341), and Lenovo (0.214), offer moderate price premiums, balancing value and reliability. Budget brands, including Vero (-0.845), Mediacom (-0.673), and Chuwi (-0.691), reduce prices, targeting affordability over premium features.

Advantages:
The study uses cleansing data, descriptive statistics, and regression modeling to provide an in-depth analysis of the factors that influence laptop prices. Model adjustment R square up to 0.8024, indicating a strong prediction accuracy, and the visual diagnosis validated the main hypothesis.

Limitations:
By excluding branding from the compact model, the effect of consumer preferences may be overlooked. Slight heteroscedasticity and high impact observations may affect the estimates. Data sets may not be fully representative of global price trends.

Ethical considerations:
The exclusion of certain variables may lead to bias in the conclusions. Transparent reporting and ensuring that the analysis is not abused, such as reinforcing unfair pricing, is critical.

Future Work:
Nonlinear models can be introduced to further explore variable interactions, and heteroscedasticity problems can be treated by weighted regression or transformation.

## References
kaggle, "Laptop Price Prediction cleaned Dataset" [Online]. Available at: https://www.kaggle.com/datasets/gyanprakashkushwaha/laptop-price-prediction-cleaned-dataset/code (Accessed: **25 November 2024**)


